---
title: "Exam 2"
execute: 
  echo: true
  warning: false
  error: false
format: 
  html:
    code-fold: true
---

# Background Processing {.unnumbered}

```{r}
# Load packages and data
library(tidyverse)
library(tidytuesdayR)
library(rnaturalearth)
library(sf)
library(ggplot2)
library(viridis)

tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```

```{r}
# Inspect & begin to tidy data
str(fc)

unique(fc$food_category)

fcc <- fc |>
  mutate(food_category = case_when(
    food_category == "Lamb & Goat" ~ "Lamb",
    food_category == "Milk - inc. cheese" ~ "Dairy",
    food_category == "Wheat and Wheat Products" ~ "Wheat",
    food_category == "Nuts inc. Peanut Butter" ~ "Nuts",
    TRUE ~ food_category  
  ))

unique(fcc$food_category)
```

# Research Question 1: Which 5 countries consume the most food?

Goal: Show the answer as a 2 column table and as a graph.
Sort countries based on consumption from largest to smallest.
Use fcc dataframe.

```{r}
top_consumers <- fcc |>
  group_by(country) |>
  summarise(total_consumption = sum(consumption, na.rm = TRUE)) |> 
  arrange(desc(total_consumption)) |>
  slice_head(n=5)

print(top_consumers)

ggplot(top_consumers, aes(x = reorder(country, -total_consumption), y = total_consumption)) +
  geom_col(fill = "steelblue") +
  labs(title = "Top 5 Countries by Food Consumtion",
       x = "Country",
       y = "Total Food Consumption (kg/person/year)")
```

# Research Question 2: For each food category what are the top 5 consumer countries?

Show the answer as a 3 column table (show the food category as the first column followed by country and consumption) and as a graph.
For each food, sort the countries based on consumption from largest to smallest.
Use the fcc dataframe.

```{r}
top_consumers_by_cat <- fcc |> 
  group_by(food_category, country) |> 
  summarise(total_consumption = sum(consumption, na.rm = TRUE), .groups = "drop") |> 
  group_by(food_category) |>  
  slice_max(order_by = total_consumption, n = 5, with_ties = FALSE) |>  
  ungroup() 


print(top_consumers_by_cat)

ggplot(top_consumers_by_cat, aes(x = reorder(country, total_consumption), y = total_consumption, fill = food_category)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ food_category, scales = "free_y") +
  coord_flip() + 
  labs(title = "Top 5 Consumer Countries for Each Food Category",
       x = "Country", y = "Total Consumption") +
  theme_minimal()

```

# Research Question 3: What does the consumption of each food look like?

Show a choropleth map for each food using the fcc dataframe.

```{r}
# Create the choropleth map with enhancements
ne_countries(returnclass = "sf") |> 
  select(name, geometry) |> 
  mutate(name = ifelse(name == "United States of America", "USA", name)) |> 
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name)) |> 
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>  # Corrected typo
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |> 
  left_join(
    fcc |> 
      select(-co2_emmission) |> 
      group_by(food_category) |> 
      mutate(consumption = (consumption - mean(consumption))/sd(consumption)), 
    by = c("name" = "country")
  ) |> 
  pivot_wider(names_from = food_category, values_from = consumption) |> 
  select(-"NA") |>  # Remove any "NA" columns
  pivot_longer(cols = c(-name, -geometry), names_to = "food_category", values_to = "consumption") |> 
  ggplot() +
  geom_sf(aes(fill = consumption), color = "gray90", size = 0.2) +  # Thinner boundary lines
  scale_fill_viridis_c(option = "C", direction = 1, na.value = "gray90") +  # Colorblind-friendly palette
  facet_wrap(~food_category) +
  labs(
    title = "Food Consumption by Country (Standardized)",
    fill = "Consumption (Z-Score)",  # Label for the color scale
    subtitle = "Each food category is visualized separately with standardized consumption values"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        strip.text = element_text(size = 10),
        panel.spacing = unit(1, "lines")) 
```
