---
title: "ica-strings"
format: html
---

### Exercise 1: Time slots {.unnumbered}

The `courses` data includes *actual* data scraped from Mac's [class schedule](https://macadmsys.macalester.edu/macssb/customPage/page/classSchedule). (Thanks to Prof Leslie Myint for the scraping code!!)

If you want to learn how to scrape data, take COMP/STAT 212, Intermediate Data Science! NOTE: For simplicity, I removed classes that had "TBA" for the `days`.

```{r}
courses <- read.csv("https://mac-stat.github.io/data/registrar.csv")

# Check it out
head(courses)
```

Use our more familiar wrangling tools to warm up.

```{r}
# Construct a table that indicates the number of classes offered in each day/time slot
# Print only the 6 most popular time slots



```



### Exercise 2: Prep the data {.unnumbered}

So that we can analyze it later, we want to wrangle the `courses` data:

-   Let's get some enrollment info:
    -   Split `avail_max` into 2 separate variables: `avail` and `max`.
    -   Use `avail` and `max` to define a new variable called `enrollment`. HINT: You'll need `as.numeric()`
-   Split the course `number` into 3 separate variables: `dept`, `number`, and `section`. HINT: You can use `separate()` to split a variable into 3, not just 2 new variables.

Store this as `courses_clean` so that you can use it later.

```{r}
```



### Exercise 3: Courses by department {.unnumbered}

Using `courses_clean`...

```{r}
# Identify the 6 departments that offered the most sections


# Identify the 6 departments with the longest average course titles

```



### Exercise 4: STAT courses {.unnumbered}

#### Part a {.unnumbered}

Get a subset of `courses_clean` that only includes courses taught by Alicia Johnson.

```{r}
```

#### Part b {.unnumbered}

Create a new dataset from `courses_clean`, named `stat`, that only includes STAT sections. In this dataset:

-   In the course names:

    -   Remove "Introduction to" from any name.
    -   Shorten "Statistical" to "Stat" where relevant.

-   Define a variable that records the `start_time` for the course.

-   Keep only the `number, name, start_time, enroll` columns.

-   The result should have 18 rows and 4 columns.

```{r}
```



### Exercise 5: More cleaning {.unnumbered}

In the next exercises, we'll dig into enrollments. Let's get the data ready for that analysis here. Make the following changes to the `courses_clean` data. Because they have different enrollment structures, and we don't want to compare apples and oranges, *remove* the following:

-   all sections in `PE` and `INTD` (interdisciplinary studies courses)

-   all music ensembles and dance practicums, i.e. all MUSI and THDA classes with numbers less than 100. HINT: `!(dept == "MUSI" & as.numeric(number) < 100)`

-   all lab sections. Be careful which variable you use here. For example, you don't want to search by "Lab" and accidentally eliminate courses with words such as "Labor".

Save the results as `enrollments` (don't overwrite `courses_clean`).

```{r}
```



### Exercise 6: Enrollment & departments {.unnumbered}

Explore enrollments by department. *You* decide what research questions to focus on. Use both visual and numerical summaries.



### Exercise 7: Enrollment & faculty {.unnumbered}

Let's now explore enrollments by instructor. In doing so, we have to be cautious of cross-listed courses that are listed under multiple different departments. Uncomment the code lines in the chunk below for an example.

::: {.callout-tip title="Commenting/Uncommenting Code"}

To comment/uncomment several lines of code at once, highlight them then click <kbd>ctrl/cmd+shift+c</kbd>.

:::

```{r}
# enrollments |>
#   filter(dept %in% c("STAT", "COMP"), number == 112, section == "01")
```

Notice that these are the exact same section! In order to not double count an instructor's enrollments, we can keep only the courses that have `distinct()` *combinations* of `days, time, instructor` values.  Uncomment the code lines in the chunk below.

```{r}
# enrollments_2 <- enrollments |> 
#   distinct(days, time, instructor, .keep_all = TRUE)

# NOTE: By default this keeps the first department alphabetically
# That's fine because we won't use this to analyze department enrollments!
# enrollments_2 |> 
#   filter(instructor == "Brianna Heggeseth", name == "Introduction to Data Science")
```

*Now*, explore enrollments by instructor. *You* decide what research questions to focus on. Use both visual and numerical summaries.

CAVEAT: The above code doesn't deal with *co-taught* courses that have more than one instructor. Thus instructors that co-taught are recorded as a pair, and their co-taught enrollments aren't added to their total enrollments. This is tough to get around with how the data were scraped as the instructor names are smushed together, not separated by a comma!

```{r}
```
